---
title: "Survey results"
format: 
  html:
    code-fold: true
    code-tools: true
    toc: true
params:
  update_data: true
  gs_name: "CPO Member Survey (Responses)"
  csv_fn: cpo_survey_2025.csv
knitr:
  opts_chunk:
    fig.path: include/fig/ 
---

## About

This page documents the procedure for cleaning and visualizing results from 
the Central Pennsylvania Observers March 2025 survey.

## Survey

Direct link: <https://docs.google.com/forms/d/e/1FAIpQLScWSF5_mUWJHatvsmPOE4APbl48tZv0FjzbnmHj8mmAiWLdPQ/viewform?usp=header>

## Prepare

First, we load the external packages (groups of R commands) that we will be using.

::: {.callout-important}

The code uses the `quietly()` function from the `purrr` package to suppress most of the feedback.

:::

```{r}
#| warning: false
library('ggplot2')
library('dplyr')

r_functions <- list.files(file.path(here::here(), "src", "R"), "\\.R$", full.names = TRUE)

purrr::map(r_functions, source) |>
  purrr::quietly()
```

## Gather

The survey data are stored in a Google Sheet.

Next, we download the data from the Google Sheet where it is collected.
Gilmore has stored his Google account credentials in a special environment file that can be accessed by the R command `Sys.getenv("GMAIL_SURVEY")`.

```{r, eval=params$update_data}
#| label: gather-data-from-google
#| message: false
#| echo: true
if (!dir.exists(file.path("include", "csv"))) {
  message("Creating missing `include/csv/`.")
  dir.create(file.path("include", "csv"))
}

if (params$update_data) {
  options(gargle_oauth_email = Sys.getenv("GMAIL_SURVEY"))
  googledrive::drive_auth()

  googledrive::drive_download(
    params$gs_name,
    path = file.path("include", "csv", params$csv_fn),
    type = "csv",
    overwrite = TRUE
  )
  message("Data updated.")
} else {
  message("Using stored data.")
}
```

## Clean

We import the downloaded data.

```{r}
results <- readr::read_csv(file = file.path("include", "csv", params$csv_fn),
                           show_col_types = FALSE)
```

### Shorten long names

The raw question values are as follows:

```{r}
names(results)
```

Then, we shorten the long names to make them more useful for data visualization and for making the data dictionary.

```{r}
renamed_cols <- results |>
  dplyr::rename("time_stamp" = "Timestamp",
                "activities_2024" = "Which of the following astronomy activities did you participate in during 2024?",
                "activities_2025" = "Which of the following astronomy activities do you plan to participate in during 2025?",
                "activities_volunteer" = "Which of the following activities would you like to get more involved with as a volunteer?",
                "astro_goals" = "Do you have any particular astronomy goals?",
                "interest_in_visual" = "Rate your level of interest in the following activities. [Visual astronomy]",
                "interest_in_astrophoto" = "Rate your level of interest in the following activities. [Astrophotography]",
                "interest_in_sci_talks" = "Rate your level of interest in the following activities. [Science-oriented talks]",
                "interest_in_social" = "Rate your level of interest in the following activities. [Social gatherings]",
                "interest_in_how_tos" = "Rate your level of interest in the following activities. [Tutorials/how-to's]",
                "interest_in_field_trips" = "Rate your level of interest in the following activities. [Field trips]",
                "interest_in_gear" = "Rate your level of interest in the following activities. [Gear and tech]",
                "interest_in_public_events" = "Rate your level of interest in the following activities. [Events for the public]",
                "club_does_well" = "What is something the club does really well?",
                "club_could_improve" = "What is something the club could do to improve?",
                "club_should_offer" = "Is there something the club should offer that we are not currently doing?",
                "anything_else" = "Anything else?",
                "respondent_name" = "Your name (totally optional)")
```

### Make data dictionary

```{r}
data_dict <- data.frame(questions = names(results), shortnames = names(renamed_cols))
```

### Split variables

Some variables have multiple responses.
These should ideally be put into separate column variables.

We'll split the activities* variables first

```{r}
split_activities <- renamed_cols |>
  dplyr::mutate(
    activities_2024_monthly_mtg = stringr::str_detect(activities_2024, "CPO monthly"),
    activities_2024_skywatch = stringr::str_detect(activities_2024, "CPO hosted Sky watch"),
    activities_2024_eclipse_personal = stringr::str_detect(activities_2024, "Personal eclipse activity"),
    activities_2024_eclipse_club = stringr::str_detect(activities_2024, "Club eclipse trip"),
    activities_2024_bfsp = stringr::str_detect(activities_2024, "Black Forest Star Party"),
    activities_2024_on_tap = stringr::str_detect(activities_2024, "Penn State Astronomy on Tap"),
    activities_2025_monthly_mtg = stringr::str_detect(activities_2025, "CPO monthly"),
    activities_2025_skywatch = stringr::str_detect(activities_2025, "CPO Sky watch"),
    activities_2025_bfsp = stringr::str_detect(activities_2025, "Black Forest Star Party"),
    activities_2025_on_tap = stringr::str_detect(activities_2025, "Penn State Astronomy on Tap")
  )
```

Then we split `astronomy_goals`.

```{r}
split_goals <- split_activities |>
  dplyr::mutate(
    more_scope = stringr::str_detect(astro_goals, "Use my telescope more"),
    buy_gear = stringr::str_detect(astro_goals, "Buy new (to me) gear"),    
    learn_sky = stringr::str_detect(astro_goals, "Learn more about the night sky"),
    astrophoto = stringr::str_detect(astro_goals, "Try astrophotography"),
    meet_folks = stringr::str_detect(astro_goals, "Meet other like-minded folks"),
    plan_trip = stringr::str_detect(astro_goals, "Plan an observing trip")
  ) |>
  dplyr::select(-c(activities_2024, activities_2025, astro_goals))
```

### Drop test cases

Joe and Chris helped test the survey back in December.
We'll drop their test responses.

```{r}
survey_cleaned <- split_goals |>
  dplyr::filter(!stringr::str_detect(time_stamp, "2024"))
```

### Convert TRUE/FALSE

```{r}
survey_cleaned <- survey_cleaned |>
  mutate(club_mtg_24 = if_else(activities_2024_monthly_mtg, "yes", "no"),
         skywatch_24 = if_else(activities_2024_skywatch, "yes", "no"),
         eclipse_personal_24 = if_else(activities_2024_eclipse_personal, "yes", "no"),
         eclipse_club_24 = if_else(activities_2024_eclipse_club, "yes", "no"),
         bfsp_24 = if_else(activities_2024_bfsp, "yes", "no"),
         on_tap_24 = if_else(activities_2024_on_tap, "yes", "no"),
         club_mtg_25 = if_else(activities_2025_monthly_mtg, "yes", "no"),
         skywatch_25 = if_else(activities_2025_skywatch, "yes", "no"),
         bfsp_25 = if_else(activities_2025_bfsp, "yes", "no"),
         on_tap_25 = if_else(activities_2025_on_tap, "yes", "no"),
         more_scope = if_else(more_scope, "yes", "no"),
         buy_gear = if_else(buy_gear, "yes", "no"),
         learn_sky = if_else(learn_sky, "yes", "no"),
         astrophoto = if_else(astrophoto, "yes", "no"),
         meet_folks = if_else(meet_folks, "yes", "no"),
         plan_trip = if_else(plan_trip, "yes", "no")
         )
```

## Visualize

As of `{r} Sys.time()`, we have had *n*=`{r} dim(survey_cleaned)[1]` responses.

### Activities 2024/Plans 2025

#### Club meetings

:::: {.columns}
::: {.column width=50%}
>`{r} data_dict$questions[2]`

```{r}
#| label: fig-attended-2024-club-meeting
#| fig-cap: "Attended a club meeting in 2024"
survey_cleaned |>
  ggplot() +
  aes(x = club_mtg_24, fill = club_mtg_24) +
  geom_bar(stat = "count", position="dodge") +
  xlab("Montly meeting") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::: {.column width=50%}
>`{r} data_dict$questions[3]`

```{r}
#| label: fig-activities_2025_monthly_mtg
#| fig-cap: "Attend a club meeting in 2025"
survey_cleaned |>
  ggplot() +
  aes(x = club_mtg_25, fill = club_mtg_25) +
  geom_bar(stat = "count", position="dodge") +
  xlab("Montly meeting") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::::

#### Skywatches

:::: {.columns}
::: {.column width=50%}
```{r}
#| label: fig-activities_2024_skywatch
#| fig-cap: "Attended a skywatch in 2024"
survey_cleaned |>
  ggplot() +
  aes(x = skywatch_24, fill = skywatch_24) +
  geom_bar(stat = "count", position="dodge") +
  xlab("Skywatches") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::: {.column width=50%}
```{r}
#| label: fig-activities_2025_skywatch
#| fig-cap: "Attend a skywatch in 2025"
survey_cleaned |>
  ggplot() +
  aes(x = skywatch_25, fill = skywatch_25) +
  geom_bar(stat = "count", position="dodge") +
  xlab("Skywatches") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::::

#### BFSP

:::: {.columns}
::: {.column width=50%}
```{r}
#| label: fig-activities_2024_bfsp
#| fig-cap: "Attended BFSP in 2024"
survey_cleaned |>
  ggplot() +
  aes(x = bfsp_24, fill = bfsp_24) +
  geom_bar(stat = "count", position="dodge") +
  xlab("BFSP") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::: {.column width=50%}
```{r}
#| label: fig-activities_2025_bfsp
#| fig-cap: "Attend BFSP in 2025"
survey_cleaned |>
  ggplot() +
  aes(x = bfsp_25, fill = bfsp_25) +
  geom_bar(stat = "count", position="dodge") +
  xlab("BFSP") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::::

#### PSU On-tap

:::: {.columns}
::: {.column width=50%}
```{r}
#| label: fig-activities_2024_on_tap
#| fig-cap: "Attended Penn State on Tap in 2024"
survey_cleaned |>
  ggplot() +
  aes(x = on_tap_24, fill = on_tap_24) +
  geom_bar(stat = "count", position="dodge") +
  xlab("Skywatch") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::: {.column width=50%}
```{r}
#| label: fig-activities_2025_on_tap
#| fig-cap: "Attend Penn State on Tap in 2025"
survey_cleaned |>
  ggplot() +
  aes(x = on_tap_25, fill = on_tap_25) +
  geom_bar(stat = "count", position="dodge") +
  xlab("Skywatch") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::::

#### 2025 Eclipse

:::: {.columns}
::: {.column width=50%}
```{r}
#| label: fig-activities_2024_eclipse_personal
#| fig-cap: "Participated in a personal eclipse activity"
survey_cleaned |>
  ggplot() +
  aes(x = eclipse_personal_24, fill = eclipse_personal_24) +
  geom_bar(stat = "count", position="dodge") +
  xlab("Personal eclipse activity") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::: {.column width=50%}
```{r}
#| label: fig-activities_2024_eclipse_club
#| fig-cap: "Participated in club eclipse trip"
survey_cleaned |>
  ggplot() +
  aes(x = eclipse_club_24, fill = eclipse_club_24) +
  geom_bar(stat = "count", position="dodge") +
  xlab("Club eclipse trip") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```
:::
::::

### Volunteering in 2025

>`{r} data_dict$questions[4]`

```{r}
survey_cleaned |>
  dplyr::select(activities_volunteer) |>
  dplyr::filter(!is.na(activities_volunteer)) |>
  knitr::kable(format = 'html')
```

### Interests

>Rate your level of interest in the following activities:

#### Visual astronomy

```{r}
#| label: fig-interest_in_visual
#| fig-cap: "Interest in visual astronomy"
survey_cleaned |>
  dplyr::mutate(interest_in_visual = factor(interest_in_visual, levels = c("Minimal interest", "Some interest", "Considerable interest"))) |>
  ggplot() +
  aes(x=interest_in_visual, fill = interest_in_visual) +
  geom_bar(stat = "count", position="dodge") +
  theme(legend.position = "none") +
  xlab("") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### Astrophotography

```{r}
#| label: fig-interest_in_astrophoto
#| fig-cap: "Interest in astrophotography"
survey_cleaned |>
  dplyr::mutate(interest_in_astrophoto = factor(interest_in_astrophoto, levels = c("Minimal interest", "Some interest", "Considerable interest"))) |>
  ggplot() +
  aes(x = interest_in_astrophoto, fill = interest_in_astrophoto) +
  geom_bar(stat = "count", position="dodge") +
  theme(legend.position = "none") +
  xlab("") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### Science talks

```{r}
#| label: fig-interest_in_sci_talks
#| fig-cap: "Interest in astrophotography"
survey_cleaned |>
  dplyr::mutate(interest_in_sci_talks = factor(interest_in_sci_talks, levels = c("Minimal interest", "Some interest", "Considerable interest"))) |>
  ggplot() +
  aes(x = interest_in_sci_talks, fill = interest_in_sci_talks) +
  geom_bar(stat = "count", position="dodge") +
  theme(legend.position = "none") +
  xlab("") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### Social gatherings

```{r}
#| label: fig-interest_in_social
#| fig-cap: "Interest in social gatherings"
survey_cleaned |>
  dplyr::mutate(interest_in_social = factor(interest_in_social, levels = c("Minimal interest", "Some interest", "Considerable interest"))) |>
  ggplot() +
  aes(x = interest_in_social, fill = interest_in_social) +
  geom_bar(stat = "count", position="dodge") +
  theme(legend.position = "none") +
  xlab("") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### How-to's

```{r}
#| label: fig-interest_in_how_tos
#| fig-cap: "Interest in how-to's"
survey_cleaned |>
  dplyr::mutate(interest_in_how_tos = factor(interest_in_how_tos, levels = c("Minimal interest", "Some interest", "Considerable interest"))) |>
  ggplot() +
  aes(x = interest_in_how_tos, fill = interest_in_how_tos) +
  geom_bar(stat = "count", position="dodge") +
  theme(legend.position = "none") +
  xlab("") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### Gear

```{r}
#| label: fig-interest_in_gear
#| fig-cap: "Interest in gear"
survey_cleaned |>
  dplyr::mutate(interest_in_gear = factor(interest_in_gear, levels = c("Minimal interest", "Some interest", "Considerable interest"))) |>
  ggplot() +
  aes(x = interest_in_gear, fill = interest_in_gear) +
  geom_bar(stat = "count", position="dodge") +
  theme(legend.position = "none") +
  xlab("") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### Field trips

```{r}
#| label: fig-interest_in_field_trips
#| fig-cap: "Interest in field trips"
survey_cleaned |>
  dplyr::mutate(interest_in_field_trips = factor(interest_in_field_trips, levels = c("Minimal interest", "Some interest", "Considerable interest"))) |>
  ggplot() +
  aes(x = interest_in_field_trips, fill = interest_in_field_trips) +
  geom_bar(stat = "count", position="dodge") +
  theme(legend.position = "none") +
  xlab("") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### Public events

```{r}
#| label: fig-interest_in_public_events
#| fig-cap: "Interest in public events"
survey_cleaned |>
  dplyr::mutate(interest_in_public_events = factor(interest_in_public_events, levels = c("Minimal interest", "Some interest", "Considerable interest"))) |>
  ggplot() +
  aes(x = interest_in_public_events, fill = interest_in_public_events) +
  geom_bar(stat = "count", position="dodge") +
  theme(legend.position = "none") +
  xlab("") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

### Goals

>`{r} data_dict$questions[5]`

#### "Use my telescope more"

```{r}
#| label: fig-more_scope
#| fig-cap: "Use my telescope more"
survey_cleaned |>
  dplyr::filter(!is.na(more_scope)) |>
  ggplot() +
  aes(x = more_scope, fill = more_scope) +
  geom_bar(stat = "count", position = "dodge") +
  xlab("Use scope more") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### "Buy new (to me) gear"

```{r}
#| label: fig-buy_gear
#| fig-cap: "Buy new gear"
survey_cleaned |>
  dplyr::filter(!is.na(buy_gear)) |>
  ggplot() +
  aes(x = buy_gear, fill = buy_gear) +
  geom_bar(stat = "count", position = "dodge") +
  xlab("Use scope more") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### "Learn more about the night sky"

```{r}
#| label: fig-learn_sky
#| fig-cap: "Learn more about the night sky"
survey_cleaned |>
  dplyr::filter(!is.na(learn_sky)) |>
  ggplot() +
  aes(x = learn_sky, fill = learn_sky) +
  geom_bar(stat = "count", position = "dodge") +
  xlab("Use scope more") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### "Try astrophotography"

```{r}
#| label: fig-astrophoto
#| fig-cap: "Learn more about astrophotography"
survey_cleaned |>
  dplyr::filter(!is.na(astrophoto)) |>
  ggplot() +
  aes(x = astrophoto, fill = astrophoto) +
  geom_bar(stat = "count", position = "dodge") +
  xlab("Astrophotography") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### "Meet other like-minded folks"

```{r}
#| label: fig-meet_folks
#| fig-cap: "Meet like-minded folks"
survey_cleaned |>
  dplyr::filter(!is.na(meet_folks)) |>
  ggplot() +
  aes(x = meet_folks, fill = meet_folks) +
  geom_bar(stat = "count", position = "dodge") +
  xlab("Meet folks") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

#### "Plan an observing trip"

```{r}
#| label: fig-plan_trip
#| fig-cap: "Plan observing trip"
survey_cleaned |>
  dplyr::filter(!is.na(plan_trip)) |>
  ggplot() +
  aes(x = plan_trip, fill = plan_trip) +
  geom_bar(stat = "count", position = "dodge") +
  xlab("Plan trip") +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10, 12))
```

### CPO as a club

#### "`{r} data_dict$questions[14]`"

```{r}
survey_cleaned |>
  dplyr::select(club_does_well) |>
  dplyr::filter(!is.na(club_does_well)) |>
  knitr::kable(format = 'html')
```

#### "`{r} data_dict$questions[15]`"

```{r}
survey_cleaned |>
  dplyr::select(club_could_improve) |>
  dplyr::filter(!is.na(club_could_improve)) |>
  knitr::kable(format = 'html')
```

#### "`{r} data_dict$questions[16]`"

```{r}
survey_cleaned |>
  dplyr::select(club_should_offer) |>
  dplyr::filter(!is.na(club_should_offer)) |>
  knitr::kable(format = 'html')
```

#### "`{r} data_dict$questions[17]`"

```{r}
survey_cleaned |>
  dplyr::select(anything_else) |>
  dplyr::filter(!is.na(anything_else)) |>
  knitr::kable(format = 'html')
```